---
title: "Homework 9 R markdown"
author: "(your name here)"
date: '`r Sys.Date()`'
output:
  html_document:
    fig_height: 4
    fig_width: 4.5
  pdf_document:
    fig_height: 4
    fig_width: 4.5
  word_document:
    fig_height: 4
    fig_width: 4.5
---


```{r, setup, include=FALSE}
require(mosaic)   # Load additional packages here 

# Some customization.  You can alter or delete as desired (if you know what you are doing).
#trellis.par.set(theme=theme.mosaic()) # change default color scheme for lattice
knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
```

#### <span style="color:Blue">**Intellectual Property:**</span>  
These problems are the intellectual property of the instructors and may not be reproduced outside of this course.

#### <span style="color:Crimson">**Due Date:**</span>  


***  
***  

##########################################################################
## Problem 1: Modeling with an Artificial Neural Network
##########################################################################

<span style="color:DarkViolet">In this problem, you will use an artificial neural network to model the type of orange juice that customers buy.  

**Important**:  Note that, for this problem, your answers should be consistent with the autograder, regardless of operating system.  If you run into any difficulties while working on a macOS, please contact your instructor with questions.



#####################################
### <span style="color:DarkViolet">Question 1</span> **<span style="color:Crimson">(4 points)</span>**:
#####################################

<span style="color:DarkViolet">**Data Set**: Load the **OJ** data set, which is in the **ISLR** library.  Set the random seed to 10.  
Use $\texttt{nnet()}$ to build an artificial neural network with 1 hidden node, to model Purchase as a function of LoyalCH, SalePriceMM, and PriceDiff.  
Enter your R code below. </span>


<span style="color:green">**Code Answer**: </span>
```{r, echo=TRUE}
library(ISLR)
library(nnet)
data(OJ)
set.seed(10)
fit = nnet(Purchase~LoyalCH+SalePriceMM+PriceDiff, data=OJ, size = 1)
```


#####################################
### <span style="color:DarkViolet">Question 2</span> **<span style="color:Crimson">(2 points)</span>**:
#####################################

<span style="color:DarkViolet">Make a plot of your neural network and upload it to the *Homework 10: Neural Network Plot* discussion.  You do not need to label the edges with their weights. </span>

<span style="color:green">**Graph Answer**  </span>: 
  (post to discussion board on D2L)
```{r,echo=FALSE}
library(NeuralNetTools)
plotnet(fit)
```



#####################################
### <span style="color:DarkViolet">Question 3</span> **<span style="color:Crimson">(3 points)</span>**:
#####################################

<span style="color:DarkViolet">In this example, the predicted response value is the probability of purchasing Minute Maid orange juice (not the probability of purchasing Citrus Hill juice).  In 2-4 sentences, explain why this makes sense based on the signs (+ or -) of the weights on the edges from **LoyalCH** to the hidden node, and from the hidden node to the response.  
**Note**: It may help to review the data dictionary using **?OJ**. </span>


<span style="color:green">**Text Answer**: The predicted response is the probability of purchasing Minute Maid orange juice. Customer brand loyalty to Citrus Hill (LoyalCH) would be expected to have a negative impact on the probability of purchasing Minute Maid. The value of LoyalCH is between 0 and 1. Larger values indicate more loyalty to Citrus Hill and produce smaller values out of the hidden node once the sigmoid function is applied. </span>

***


#####################################
### <span style="color:DarkViolet">Question 4</span> **<span style="color:Crimson">(1 point)</span>**:
#####################################

<span style="color:DarkViolet">What is the predicted probability that the first person in the data set will purchase Minute Maid? </span>

<span style="color:green">**Numeric Answer**  0.3331203</span> 
**<span style="color:red">(AUTOGRADED on D2L)</span>**:  
```{r,echo=FALSE}
predict(fit, OJ[1,])
```

#####################################
### <span style="color:DarkViolet">Question 5</span> **<span style="color:Crimson">(2 points)</span>**:
#####################################

<span style="color:DarkViolet">Compute $\sigma(z)$, the output of the hidden node, for the first person in the data set. </span>

<span style="color:green">**Numeric Answer**  0.5046999</span> 
**<span style="color:red">(AUTOGRADED on D2L)</span>**:  
```{r,echo=FALSE}
OJ[1,c("LoyalCH","SalePriceMM","PriceDiff")]
z=-1.63 + 4.22*OJ[1,"LoyalCH"] - 0.52*OJ[1,"SalePriceMM"] + 2.39*OJ[1,"PriceDiff"]
1/(1+exp(-z))
```

#####################################
### <span style="color:DarkViolet">Question 6</span> **<span style="color:Crimson">(2 points)</span>**:
#####################################

<span style="color:DarkViolet">If we use a probability threshold of .5 to classify predicted purchases, what is the classification error rate for this data set? </span>

<span style="color:green">**Numeric Answer**  </span> 
**<span style="color:red">(AUTOGRADED on D2L)</span>**:  
```{r,echo=FALSE}

```

***

#####################################
### <span style="color:DarkViolet">Question 7</span> **<span style="color:Crimson">(1 point)</span>**:
#####################################

<span style="color:DarkViolet">Suppose we classify predicted purchases as "MM"" if the probability of purchasing Minute Maid is > .9, as "CH" if the probability of purchasing Minute Maid is < .1, and NA otherwise. What is the classification error rate among purchases for which we make a (non-NA) prediction? </span>


<span style="color:green">**Numeric Answer**  </span> 
**<span style="color:red">(AUTOGRADED on D2L)</span>**:  
```{r,echo=FALSE}
```

#####################################
### <span style="color:DarkViolet">Question 8</span> **<span style="color:Crimson">(3 points)</span>**:
#####################################

<span style="color:DarkViolet">Write the R code you used to answer the previous question. </span>

<span style="color:green">**Text Answer**: </span>


#####################################
### <span style="color:DarkViolet">Question 9</span> **<span style="color:Crimson">(2 points)</span>**:
#####################################

<span style="color:DarkViolet">If we use a probability threshold of .9 as in the previous two questions, for how many purchases do we fail to make a (non-NA) prediction? </span>

<span style="color:green">**Numeric Answer**  </span> 
**<span style="color:red">(AUTOGRADED on D2L)</span>**:  
```{r,echo=FALSE}
```


#####################################
### <span style="color:DarkViolet">Question 10</span> **<span style="color:Crimson">(2 points)</span>**:
#####################################

<span style="color:DarkViolet">View the Lek profile of the model. Which of the following accurately describe the relationship among the variables? Select all that apply. </span>

<span style="color:green">**Multiple SELECT Answer** </span>
  **<span style="color:red">(AUTOGRADED on D2L)</span>**:  
  
The association between LoyalCH and Purchase is stronger for customers with SalePriceMM in the 80th percentile than in the 60th percentile.

The association between SalePriceMM and Purchase is strongest for customers with LoyalCH in the 40th percentile.

LoyalCH and Purchase are negatively associated.

PriceDiff and Purchase are positively associated.


***
***

##########################################################################
## Problem 2: Using an Artificial Neural Network to Model Salaries
##########################################################################

<span style="color:DarkViolet">In this problem, you will use an artificial neural network to model the salaries of baseball players. 
**Important**:  Note that, for this problem, your answers may be different if you are working on a macOS; thus, there are no numeric autograded parts.</span>

#####################################
### <span style="color:DarkViolet">Question 11</span> **<span style="color:Crimson">(3 points)</span>**:
#####################################

<span style="color:DarkViolet">**Data Set**: Load the **Hitters** data set in the **ISLR** package.
Remove any rows with missing Salary data.  
Create new variables as follows, adding to the data frame in the order listed:

1. **League01**, which equals 0 if **League** = "A" and equals 1 if **League** = "N".   
2. **Division01**, which equals 0 if **Division** = "E" and equals 1 if **Division** = "W".  
3. **NewLeague01**, which equals 0 if **NewLeague** = "A" and equals 1 if **NewLeague** = "N".  

<span style="color:DarkViolet">*Remove* the old variables (**League**, **NewLeague**, and **Division**) from the data frame.  
Enter your R code below. </span>

<span style="color:green">**Code Answer**: </span>
```{r, echo=TRUE}
```


#####################################
### <span style="color:DarkViolet">Question 12</span> **<span style="color:Crimson">(5 points)</span>**:
#####################################

<span style="color:DarkViolet">We will use $\texttt{nnet()}$ to fit an artificial neural network with 10 hidden nodes to model **Salary** as a function of all other variables in the data set.  Set the random seed equal to 10.  Let 
    $\texttt{decayRate = seq(.1, 3, by = .1)}$  
Use 10-fold cross-validation to choose the best decay rate, $\lambda$.  Use 10 hidden nodes and a linear output function.  To ensure convergence, use $\texttt{maxit = 1000}$.  
Within each fold (or iteration of the cross-validation loop), you should create a new data frame that contains the training data, and then standardize all of the variables in it (including Salary) to have mean 0 and standard deviation 1.

*  Standardizing the response variable is convenient for treating all of the variables in the same way, reducing the influence of outliers, and avoiding huge values for the weights and MSE.

<span style="color:DarkViolet">In each fold, you should also create a new data frame that contains the *validation* data, and then standardize all of the variables in it based on the mean and standard deviation you used to standardize the training data.

* Depending on how you standardize the validation data, you may get an error message when trying to access 
$\texttt{Hitters.valid\$Salary}$.  If so, refer to this variable by its column number instead; e.g. $\texttt{Hitters.valid[ ,17]}$.
This analysis may take a few minutes to run in R.

<span style="color:DarkViolet">Hint to help you check your work:  The MSE when using $\lambda $ = .1 should be .6371 (**if** you are working on a PC).
Enter your R code below. </span>


<span style="color:green">**Code Answer**: </span>
```{r, echo=TRUE}
```


***


#####################################
### <span style="color:DarkViolet">Question 13</span> **<span style="color:Crimson">(3 points)</span>**:
#####################################

<span style="color:DarkViolet">What value of $\lambda$ minimizes the MSE? *Discuss* whether any other values of $\lambda$ appear reasonable, based on the observed values of MSE.</span>


<span style="color:green">**Text Answer**: </span>



***

#####################################
### <span style="color:DarkViolet">Question 14</span> **<span style="color:Crimson">(1 point)</span>**:
#####################################

<span style="color:DarkViolet">Set the random seed to 10 again.  Standardize the entire data set, and use the best value of $\lambda$ from the previous problem to fit a neural network to the entire data set.  Use 10 hidden nodes and a linear output function.  Use a maximum of 300 iterations; does the model converge? </span>

<span style="color:green">**Multiple choice Answer** </span>
  **<span style="color:red">(AUTOGRADED on D2L)</span>**:  one of  
Yes,  
No

#####################################
### <span style="color:DarkViolet">Question 15</span> **<span style="color:Crimson">(2 points)</span>**:
#####################################

<span style="color:DarkViolet">Apply Garson's algorithm to the model from the previous question. Which _two_ (standardized) variables are most important in predicting (standardized) **Salary**? </span>


<span style="color:green">**Text Answer**: </span>


#####################################
### <span style="color:DarkViolet">Question 16</span> **<span style="color:Crimson">(2 points)</span>**:
#####################################

<span style="color:DarkViolet">If your goal was to optimize the performance of this model, what would you try next? Why? Explain in 1-2 sentences. </span>


<span style="color:green">**Text Answer**: </span>




#####################################
### <span style="color:DarkViolet">Question 17</span> **<span style="color:Crimson">(1 point)</span>**:
#####################################

<span style="color:DarkViolet"> What operating system did you use to run the code for this Homework?
**Note**: All options for this question are correct; you get the point automatically for providing feedback.</span>


<span style="color:green">**Multiple choice Answer** </span>
  **<span style="color:red">(AUTOGRADED on D2L)</span>**:  one of  
	
Windows (PC),  
Virtual Lab (Windows/PC),  
Mac,  
Linux or other
