---
title: "Homework 8 R markdown"
author: "(your name here)"
date: '`r Sys.Date()`'
output:
  html_document:
    fig_height: 4
    fig_width: 4.5
  pdf_document:
    fig_height: 4
    fig_width: 4.5
  word_document:
    fig_height: 4
    fig_width: 4.5
---


```{r, setup, include=FALSE}
require(mosaic)   # Load additional packages here 

# Some customization.  You can alter or delete as desired (if you know what you are doing).
trellis.par.set(theme=theme.mosaic()) # change default color scheme for lattice
knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
```

#### <span style="color:Blue">**Intellectual Property:**</span>  
These problems are the intellectual property of the instructors and may not be reproduced outside of this course.

#### <span style="color:Crimson">**Due Date:**</span>  
Tuesday, November 7, 2017 at 11:59 PM 

***  
***  


##########################################################################
## Problem 1: Using Support Vectors for Classification
##########################################################################

<span style="color:DarkViolet">In this problem, you will use a support vector classifier to categorize different species of oak trees to different regions (California vs. the Atlantic coast).  
**Data Set**: Download the *Oak_log.csv* data set and load it into R.  </span>



#####################################
### <span style="color:DarkViolet">Question 1</span> **<span style="color:Crimson">(3 points)</span>**:
#####################################

<span style="color:DarkViolet">Make a scatterplot that shows the region, standardized log(acorn size), and standardized log(range) all in the same plot.  Include a legend.  
Enter your R code below.  </span>


<span style="color:green">**Code Answer**: </span>
```{r, echo=TRUE}
oaklog = read.csv("Oak_log.csv") 
attach(oaklog)
my.col = rep("red",39)
my.col[which(Region=="Atlantic")] = "blue" 
my.pch = rep(21,39)
my.pch[which(Region=="Atlantic")] = 4

plot(scale(logSize),scale(logRange),col=my.col,pch=my.pch, xlab = "Standardized Log Size", ylab = "Standardized Log Range")
legend("topleft", legend = c("California", "Atlantic"), pch = c(21, 4), col = c("red","blue"))

```


#####################################
### <span style="color:DarkViolet">Question 2</span> **<span style="color:Crimson">(2 points)</span>**:
#####################################

<span style="color:DarkViolet">Comment on the appropriateness of a support vector classifier to categorize the regions of species of oak trees. </span>


<span style="color:green">**Text Answer**: A support vector classifier would work well with the oak log dataset, because the classes Atlantic and California are well separated. </span>


#####################################
### <span style="color:DarkViolet">Question 3</span> **<span style="color:Crimson">(2 points)</span>**:
#####################################

<span style="color:DarkViolet">Use a support vector classifier (a SVM with a linear kernel) to categorize the data points.  Use a cost of 1.  
Enter your R code below. </span>


<span style="color:green">**Code Answer**: </span>
```{r, echo=TRUE}
library(e1071)
svmfit = svm(Region~logSize+logRange, data = oaklog, kernel = "linear", cost = 1)
```


#####################################
### <span style="color:DarkViolet">Question 4</span> **<span style="color:Crimson">(2 points)</span>**:
#####################################

<span style="color:DarkViolet">Treating the entire data set as the training data, how many Atlantic species are incorrectly classified as California species?</span>

<span style="color:green">**Numeric Answer**  0.07142857</span> 
**<span style="color:red">(AUTOGRADED on D2L)</span>**:  
```{r}
summary(svmfit)

#confusion matrix
table(svmfit$fitted,Region)
2/(2+26) #incorrect atlantic
0/(0+11) #incorrect california
table(Region)
```

#####################################
### <span style="color:DarkViolet">Question 5</span> **<span style="color:Crimson">(1 point)</span>**:
#####################################

<span style="color:DarkViolet">How many California species are incorrectly classified as Atlantic species?</span>

<span style="color:green">**Numeric Answer**  0</span> 
**<span style="color:red">(AUTOGRADED on D2L)</span>**:  


#####################################
### <span style="color:DarkViolet">Question 6</span> **<span style="color:Crimson">(3 points)</span>**:
#####################################

<span style="color:DarkViolet">Plot the optimal hyperplane (line) with its margins on the same plot from the first question in this problem.  Upload your plot to the *Homework 8: Optimal Hyperplane* discussion board.   </span>

<span style="color:green">**Graph Answer**  </span>: 
  (post to discussion board on D2L)
```{r,echo=FALSE}
w = colSums(svmfit$coefs[,1] * svmfit$SV)
b = svmfit$rho
# Optimal line
abline(b/w[2],-w[1]/w[2]) # y-int, slope
# Margin lines
abline((b-1)/w[2],-w[1]/w[2], lty=2) # y-int, slope; 
# as cost decreases, w[2] tends to decrease, resulting in a larger margin
abline((b+1)/w[2],-w[1]/w[2], lty=2) 
```



***
***

##########################################################################
## Problem 2:  Using Support Vector Machines for Classification
##########################################################################

<span style="color:DarkViolet">In this problem, you will use support vector machines with different kernels to classify the gas mileage of cars.  
**Data Set**: Load the **Auto** data set, which is in the **ISLR** library.</span>




#####################################
### <span style="color:DarkViolet">Question 7</span> **<span style="color:Crimson">(2 points)</span>**:
#####################################

<span style="color:DarkViolet">Create a binary variable that equals 1 for cars with gas mileage above the median and a 0 for cars with gas mileage below the median.  Tell R to treat it as a factor.  
Tell R to treat the origin variable as a factor.  
Enter your R code below. </span>


<span style="color:green">**Code Answer**: </span>
```{r, echo=TRUE}
library(ISLR)
data("Auto")
Auto$origin = as.factor(Auto$origin) 
mpg_median = median(Auto$mpg)
Auto$mpg_above_median = rep(0,392)
Auto$mpg_above_median[which(Auto$mpg > mpg_median)] = 1 
Auto$mpg_above_median <- factor(Auto$mpg_above_median,levels = c(0,1),labels = c("Below", "Above"))
Auto = Auto[,-c(1,9)]
attach(Auto)
```

#####################################
### <span style="color:DarkViolet">Question 8</span> **<span style="color:Crimson">(2 points)</span>**:
#####################################

<span style="color:DarkViolet">Use a support vector classifier to model the binary variable gas mileage as a function of the other variables in Auto.
 
* <span style="color:DarkViolet"> Do not include the continuous variable gas mileage.</span>
* <span style="color:DarkViolet"> Do not include the car’s name.</span>
* <span style="color:DarkViolet"> Use a cost of 1.</span> 

<span style="color:DarkViolet"> Enter your R code below.  </span>


<span style="color:green">**Code Answer**: </span>
```{r, echo=TRUE}
library(e1071)
model=mpg_above_median~.
svmfit = svm(model,data=Auto,kernel="linear",cost=1)
```



#####################################
### <span style="color:DarkViolet">Question 9</span> **<span style="color:Crimson">(3 points)</span>**:
#####################################

<span style="color:DarkViolet">Set your random seed equal to 9.  Use a built-in function to perform cross-validation to compare different values of cost:  .001, .01, .1, 1, 5, 10, 100.  
Enter your R code below.  </span>


<span style="color:green">**Code Answer**: </span>
```{r, echo=TRUE}
set.seed(9)
model=mpg_above_median~.
tune.out=tune(svm,model,data=Auto,kernel="linear",ranges=list(cost=c(.001, .01, .1, 1, 5, 10, 100)))

#dont need this in quiz
best.mod = tune.out$best.model


summary(best.mod)

summary(tune.out)
summary(tune.out)$best.performance
```

#####################################
### <span style="color:DarkViolet">Question 10</span> **<span style="color:Crimson">(1 point)</span>**:
#####################################

<span style="color:DarkViolet">Which cost gave the lowest cross-validation error? </span>


<span style="color:green">**Numeric Answer**  10</span> 
**<span style="color:red">(AUTOGRADED on D2L)</span>**: 


#####################################
### <span style="color:DarkViolet">Question 11</span> **<span style="color:Crimson">(1 point)</span>**:
#####################################

<span style="color:DarkViolet">What was the cross-validation error of the best model?</span>


<span style="color:green">**Numeric Answer**  0.07628205</span> 
**<span style="color:red">(AUTOGRADED on D2L)</span>**: 
  


***

#####################################
### <span style="color:DarkViolet">Question 12</span> **<span style="color:Crimson">(2 points)</span>**:
#####################################

<span style="color:DarkViolet">Perform cross-validation to compare different values of cost and gamma for a radial model. Use the same values of cost as listed previously: .001, .01, .1, 1, 5, 10, 100. Use gamma =0.5, 1, 2, 3, and 4.  
Enter your R code below.  </span>


<span style="color:green">**Code Answer**: </span>
```{r, echo=TRUE}
#tune.out=tune(svm,y~x,kernel="radial",ranges=list(cost=c(.001,.01,.1,1,5,10,100),gamma=c(.5,1,2,3,4)), type="C-classification")

#set.seed(9)
model=mpg_above_median~.
tune.out=tune(svm,model,data=Auto,kernel="radial",ranges=list(cost=c(.001, .01, .1, 1, 5, 10, 100),gamma=c(.5,1,2,3,4)), type="C-classification")

#don't need this
best.mod = tune.out$best.model
summary(best.mod)

summary(tune.out)
summary(tune.out)$best.performance
```


#####################################
### <span style="color:DarkViolet">Questions 13-14 </span> **<span style="color:Crimson">(2 points, 1 each)</span>**:
#####################################


<span style="color:DarkViolet">Which combination of parameters gave the lowest cross-validation error?  </span>

<span style="color:DarkViolet">Cost:  </span>  
<span style="color:green">**Numeric Answer**  1</span> 
**<span style="color:red">(AUTOGRADED on D2L)</span>**:  


<span style="color:DarkViolet">Gamma: </span>  
<span style="color:green">**Numeric Answer**  3</span> 
**<span style="color:red">(AUTOGRADED on D2L)</span>**:  


#####################################
### <span style="color:DarkViolet">Question 15 </span> **<span style="color:Crimson">(1 point)</span>**:
#####################################

<span style="color:DarkViolet">What was the cross-validation error of the best model? </span>  

<span style="color:green">**Numeric Answer**  0.07384615</span> 
**<span style="color:red">(AUTOGRADED on D2L)</span>**:  


#####################################
### <span style="color:DarkViolet">Question 16</span> **<span style="color:Crimson">(3 points)</span>**:
#####################################

<span style="color:DarkViolet">Use the best model to predict whether the following car would have high or low gas mileage:  
**1977 Chrysler Sunbeam**  
**Cylinders**: 4  
**Engine displacement**:  132.5 cubic inches  
**Horsepower**:  155  
**Weight**:  2,910 lbs  
**Acceleration**:  8.3 seconds  
**Origin**:  American (1)  
Enter your R code below.  </span>

<span style="color:green">**Code Answer**: </span>
```{r, echo=TRUE}
#summary(Auto)

#best.mod = tune.out$best.model
newData = data.frame(cylinders=4,displacement=132.5,horsepower=155,weight=2910,acceleration=8.3,year=77,origin=1)
newData$origin = factor(newData$origin,levels=c(1,2,3))
predict(best.mod, newdata = newData)
```


#####################################
### <span style="color:DarkViolet">Question 17</span> **<span style="color:Crimson">(1 point)</span>**:
#####################################

<span style="color:DarkViolet">What is your prediction?</span>


<span style="color:green">**Multiple choice Answer** </span>
  **<span style="color:red">(AUTOGRADED on D2L)</span>**:  one of  
Low,  
->High  
